String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-12 2023-1-19/sections/algorithms.tex
experiments
revise syntax in algorithms
space adjustment
space ajustment
section 5 and 6

The goal of this section is to show how we solve RECL constraints by utilizing CEFA. 
At first, we show how to reduce the satisfiability of RECL constraints into a decision problem for CEFA defined in the sequel. Then we propose algorithms to solve the decision problem.

\begin{definition}[Nonemptiness of CEFA w.r.t. an LIA formula\footnote{The decision problem was called satisfiability of an LIA formula w.r.t. CEFA in \cite{atva2020}.}, abbreviated as $\cefadec$]
Suppose that $x_1, \cdots, x_n$ are mutually distinct string variables, $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are nonempty sets of CEFA over the same alphabet $\Sigma$ with $\Lambda_{x_i} = \{\aut_{i, 1}, \cdots, \aut_{i, l_i}\}$ for every $i \in [n]$, moreover, the sets of registers $R_{\aut_{1, 1}}, \cdots, R_{\aut_{1, l_1}}, \cdots, R_{\aut_{n, 1}}, \cdots, R_{\aut_{n, l_n}}$ are mutually disjoint, and $\varphi$ is an LIA formula whose free variables are from $R' = \bigcup \limits_{i \in [n], j \in [l_i]} R_{\aut_{i, j}}$. Then  the CEFA in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are said to be nonempty w.r.t. $\varphi$ if there are an assignment $\theta: \{x_1, \cdots, x_n\} \rightarrow \Sigma^*$ and vectors 
$\myvec{v_{1,1}}, \cdots, \myvec{v_{1, l_1}}, \cdots, \myvec{v_{n, 1}}, \cdots, \myvec{v_{n, l_n}}$ such that $(\theta(x_i); \myvec{v_{i,j}}) \in \Lang(\aut_{i, j})$ for every $i \in [n], j \in [l_i]$,
% \in \aut_{i,1}(\eta(x_1)), \cdots, \myvec{v_{i, l_i}} \in \aut_{i,l_i}(\eta(x_i))$ for every $i \in [n]$  
%\cdots, \myvec{v_{n,1}} \in \aut_{n,1}(w_n), \cdots, \myvec{v_{n, l_n}} \in \aut_{n, l_n}(w_n)$ 
and $\varphi[\myvec{v_{1,1}}/R_{\aut_{1,1}}, \cdots, \myvec{v_{n, l_n}}/R_{\aut_{n, l_n}}]$ is true. 
\end{definition}

\begin{proposition}[\cite{atva2020}]\label{prop-cefadec}
$\cefadec$ is PSPACE-complete. 
\end{proposition}
The decision procedure for proving the upper bound in Proposition~\ref{prop-cefadec} in \cite{atva2020}  was not implemented as a matter of fact. Instead, the symbolic model checker nuXmv was used to solve $\cefadec$ in \cite{atva2020}. We do not rely on nuXmv in this work and will propose an efficient algorithm for solving $\cefadec$ in Section~\ref{subsec:cefadec}.

@@

\subsection{From $\reclsat$ to $\cefadec$} \label{subsec:regex2cefa}

Let $\varphi$ be an RECL constraint and $x_1, \cdots, x_n$ be an enumeration of the string variables occurring in $\varphi$. Moreover, let $\varphi \equiv \varphi_1 \wedge \varphi_2$ such that $\varphi_1$ is a conjunction of regular membership constraints of $\varphi$, and $\varphi_2$ is a conjunction of length constraints of $\varphi$.
We shall reduce the satisfiability of $\varphi$ to an instance of $\cefadec$.

@@

\subsection{Solving $\cefadec$} \label{subsec:cefadec}

>>>>

The goal of this section is to show how we solve RECL constraints by utilizing CEFA. 
At first, we reduce the satisfiability of RECL constraints into a decision problem for CEFA defined in the sequel. Then we propose an algorithm to solve the decision problem.

\begin{definition}[Nonemptiness of CEFA w.r.t. an LIA formula\footnote{The decision problem was called satisfiability of an LIA formula w.r.t. CEFA in \cite{atva2020}.}, abbreviated as $\cefadec$]
Suppose that $x_1, \cdots, x_n$ are string variables, $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are nonempty sets of CEFA over the alphabet $\Sigma$ with $\Lambda_{x_i} = \{\aut_{i, 1}, \cdots, \aut_{i, l_i}\}$ for every $i \in [n]$, and the sets of registers $R_{\aut_{1, 1}}, \cdots, R_{\aut_{1, l_1}}, \cdots, R_{\aut_{n, 1}}, \cdots, R_{\aut_{n, l_n}}$ are mutually disjoint, moreover, $\varphi$ is an LIA formula whose free variables are from $R' = \bigcup \limits_{i \in [n], j \in [l_i]} R_{\aut_{i, j}}$. Then  the CEFA in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ are said to be nonempty w.r.t. $\varphi$ if there are an assignment $\theta: \{x_1, \cdots, x_n\} \rightarrow \Sigma^*$ and vectors $\myvec{v_{i,j}}$ for $i \in [n], j \in [l_i]$
%$\myvec{v_{1,1}}, \cdots, \myvec{v_{1, l_1}}, \cdots, \myvec{v_{n, 1}}, \cdots, \myvec{v_{n, l_n}}$ 
such that $(\theta(x_i); \myvec{v_{i,j}}) \in \Lang(\aut_{i, j})$ for every $i \in [n], j \in [l_i]$,
% \in \aut_{i,1}(\eta(x_1)), \cdots, \myvec{v_{i, l_i}} \in \aut_{i,l_i}(\eta(x_i))$ for every $i \in [n]$  
%\cdots, \myvec{v_{n,1}} \in \aut_{n,1}(w_n), \cdots, \myvec{v_{n, l_n}} \in \aut_{n, l_n}(w_n)$ 
and 
$\varphi[(\myvec{v_{i,j}}/R_{\aut_{i,j}})_{i \in [n], j \in [l_i]}]$ is true. 
%$\varphi[\myvec{v_{1,1}}/R_{\aut_{1,1}}, \cdots, \myvec{v_{n, l_n}}/R_{\aut_{n, l_n}}]$ is true. 
\end{definition}

\begin{proposition}[\cite{atva2020}]\label{prop-cefadec}
$\cefadec$ is PSPACE-complete. 
\end{proposition}
The decision procedure in \cite{atva2020} was only used to prove the upper bound in Proposition~\ref{prop-cefadec} and not implemented as a matter of fact. Instead, the symbolic model checker nuXmv was used to solve $\cefadec$ in \cite{atva2020}. We do not rely on nuXmv in this work and shall propose an efficient algorithm for solving $\cefadec$ in Section~\ref{subsec:cefadec}.

@@

\vspace{-2mm}
\subsection{From $\reclsat$ to $\cefadec$} \label{subsec:regex2cefa}

Let $\varphi$ be a RECL constraint and $x_1, \cdots, x_n$ be an enumeration of the string variables occurring in $\varphi$. Moreover, let $\varphi \equiv \varphi_1 \wedge \varphi_2$ such that $\varphi_1$ is a conjunction of regular membership constraints of $\varphi$, and $\varphi_2$ is a conjunction of length constraints of $\varphi$.
We shall reduce the satisfiability of $\varphi$ to an instance of $\cefadec$.

@@

\vspace{-2mm}
\subsection{Solving $\cefadec$} \label{subsec:cefadec}

### 2023-05-12 2023-1-19/sections/cefa.tex
space ajustment
section 5 and 6

%
%In~\cite{atva2020}, the cost function is defined as a function $\eta: \Sigma \rightarrow \mathbb{N}$. In this paper, we define the cost function as an integer vector whose elements are the incremental value of registers. Furthermore, we add a new linear integer arithmetic constraint $\theta$ to each final state of the CEFA, which restricts the value of registers. Two types of definitions have the same expressive ability on the $SAT_{CL}$ problem.
\begin{definition}[Cost-Enriched Finite Automaton]
  A cost-enriched finite automaton $\aut$ is a tuple $(R, Q, \Sigma, \delta, I, F)$ where
  \begin{itemize}
  \item $R = \{r_1, \cdots, r_k\}$ is a finite set of registers, 
    \item $Q, \Sigma, I, F$ are as in the definition of NFA,
%    \item $R = (r_1\cdots r_n)$ is a vector of mutually distinct cost registers,
    \item $\delta \subseteq Q \times \Sigma \times Q \times \Int^R$ is a transition relation, where $\Int^R$ denotes the updates on the values of registers.
%  set whose elements are tuples $(q, c, q', \myvec{v})$ where $q, q'$ are states of $Q$, $c$ is a letter in alphabet $\Sigma\cup\{\epsilon\}$ and $\myvec{v}$ is the cost update function for registers, which is an integer vector whose $i$-th element is the incremental value of register $r_i$.  $(q, a, q', \myvec{v})$ is written as $q\xrightarrow[\myvec{v}]{a} q'$ for readability.
    \item $\alpha \in \Phi(R)$ is an LIA formula specifying an accepting condition.
%    a linear integer arithmetic constraint function on final states. $\theta$ is called \emph{accepting condition}.
  \end{itemize}
\end{definition}
%The definition of CEFA above is slightly different from the definition in~\cite{atva2020} in the sense that accepting conditions on register values
%are attached to final states.

For readability, from now on, we assume a linear order on $R$ and write $R$ as a vector $\myvec{r} = (r_1, \cdots, r_k)$. Accordingly, we write an element of $\Int^R$ as a vector $(v_1, \cdots, v_k)$, where for each $i \in [k]$, $v_i$ is the update on the value of $r_i$. We also write a transition $(q, a, q', \vec{v}) \in \delta$ as $q \xrightarrow[\vec{v}]{a} q'$.

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{configuration} of $\aut$ is a pair $(q, \vec{v})$ where $q \in Q$ and $\vec{v}$ is a vector denoting the values of registers.  
%An \emph{initial configuration} of $\aut$ is $(q_0, \vec{0})$ with $q_0 \in I$, where the value of each register is zero. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$. Note that we assume that the initial values of all registers are zero and $\sum \limits_{j \in [n]} \myvec{v_j}$ is the vector of register values after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

We would like to remark that the accepting conditions $\alpha$ in CEFA are defined in a \emph{global} fashion in the sense that the accepting condition does not distinguish final states. This technical choice is made so that the determinization and minimization of NFA can be utilized to reduce the size of CEFA in the next section.

@@

For convenience, for a CEFA $\aut$, we use $R_\aut$ to denote the set of registers of $\aut$.

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$, is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states. Two fresh auxiliary registers, say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ or from $\aut_2$. Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$, where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
\item $\delta'$ is the union of $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$, $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$, $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$, and $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$, where the last two components of the vectors denote the updates of $r'_1$ and $r'_2$,

@@

The \emph{concatenation} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \concat \aut_2$, is defined similarly as that of NFA, that is, a tuple $(Q', \Sigma, \delta', I', F', \alpha')$, where $Q' = Q_1 \cup Q_2$, $I' = I_1$, $\alpha' = \alpha_1 \wedge \alpha_2$, $\delta' = \{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0})) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\} \cup \{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2})) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\} \cup \{(q_1, a, q_2, \vec{v_2}) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2, \vec{v_2}) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$. From the construction, 
$$\Lang(\aut_1 \concat \aut_2) = \{(w_1w_2; \myvec{v}) \mid (w_1; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1), (w_2; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$

>>>>

%
%In~\cite{atva2020}, the cost function is defined as a function $\eta: \Sigma \rightarrow \mathbb{N}$. In this paper, we define the cost function as an integer vector whose elements are the incremental value of registers. Furthermore, we add a new linear integer arithmetic constraint $\theta$ to each final state of the CEFA, which restricts the value of registers. Two types of definitions have the same expressive ability on the $SAT_{CL}$ problem.
\begin{definition}[Cost-Enriched Finite Automaton]
  A cost-enriched finite automaton $\aut$ is a tuple $(R, Q, \Sigma, \delta, I, F)$ where
  \begin{itemize}
  \item $R = \{r_1, \cdots, r_k\}$ is a finite set of registers, 
    \item $Q, \Sigma, I, F$ are as in the definition of NFA,
%    \item $R = (r_1\cdots r_n)$ is a vector of mutually distinct cost registers,
    \item $\delta \subseteq Q \times \Sigma \times Q \times \Int^R$ is a transition relation, where $\Int^R$ denotes the updates on the values of registers.
%  set whose elements are tuples $(q, c, q', \myvec{v})$ where $q, q'$ are states of $Q$, $c$ is a letter in alphabet $\Sigma\cup\{\epsilon\}$ and $\myvec{v}$ is the cost update function for registers, which is an integer vector whose $i$-th element is the incremental value of register $r_i$.  $(q, a, q', \myvec{v})$ is written as $q\xrightarrow[\myvec{v}]{a} q'$ for readability.
    \item $\alpha \in \Phi(R)$ is an LIA formula specifying an accepting condition.
%    a linear integer arithmetic constraint function on final states. $\theta$ is called \emph{accepting condition}.
  \end{itemize}
\end{definition}
For a CEFA $\aut$, we use $R_\aut$ to denote the set of registers of $\aut$. 
%The definition of CEFA above is slightly different from the definition in~\cite{atva2020} in the sense that accepting conditions on register values
%are attached to final states.

For convenience, from now on, we assume a linear order on $R$ and write $R$ as a vector $\myvec{r} = (r_1, \cdots, r_k)$. Accordingly, we write an element of $\Int^R$ as a vector $(v_1, \cdots, v_k)$, where for each $i \in [k]$, $v_i$ is the update on the value of $r_i$. We also write a transition $(q, a, q', \vec{v}) \in \delta$ as $q \xrightarrow[\vec{v}]{a} q'$.

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
%A \emph{configuration} of $\aut$ is a pair $(q, \vec{v})$ where $q \in Q$ and $\vec{v}$ is a vector denoting the values of registers.  
%An \emph{initial configuration} of $\aut$ is $(q_0, \vec{0})$ with $q_0 \in I$, where the value of each register is zero. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$. Note that we assume that the initial values of all registers are zero and $\sum \limits_{j \in [n]} \myvec{v_j}$ is the vector of register values after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

We would like to remark that the aforementioned definition of CEFA is slightly different from that in \cite{atva2020}, where CEFA did not include accepting conditions.
Moreover, the accepting conditions $\alpha$ in CEFA are defined in a \emph{global} fashion in the sense that the accepting condition does not distinguish final states. This technical choice is made so that the determinization and minimization of NFA can be utilized to reduce the size of CEFA in the next section.

@@

@@

The \emph{union} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cup \aut_2$, is defined in a slightly more involved manner than NFA, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states. Two fresh auxiliary registers, say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ or from $\aut_2$. Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$, where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
\item $\delta'$ is the union of $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$, $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$, $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$, and $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$, where $\myvec{0}$ in $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector of updates on $R_2$ and $1,0$, the last two components of the vector, denote the updates on $r'_1$ and $r'_2$, similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

@@

The \emph{concatenation} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \concat \aut_2$, is defined similarly as that of NFA, that is, a tuple $(Q', \Sigma, \delta', I', F', \alpha')$, where $Q' = Q_1 \cup Q_2$, $I' = I_1$, $\alpha' = \alpha_1 \wedge \alpha_2$, $\delta' = \{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0})) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\} \cup \{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2})) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\} \cup \{(q_1, a, q_2, (\myvec{0}, \vec{v_2})) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2, \vec{v_2}) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$. From the construction, 
$$\Lang(\aut_1 \concat \aut_2) = \{(w_1w_2; \myvec{v}) \mid (w_1; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1), (w_2; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$

### 2023-05-12 2023-1-19/sections/overview.tex
intro
section 5 and 6

We first construct an automaton for the regular expression $(\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}}$. We introduce three registers, say $r_1, r_2, r_3$, to represent the three counting operators.  
%Moreover, we introduce another register, say $r_0$, to denote the length of strings. 
Then we first construct a nondeterministic finite automaton (NFA) for $(\Sigma \setminus a)^* (\Sigma \setminus b)^* (\Sigma \setminus c)^*$, then add the updates of registers to transitions, resulting in a so-called cost-enriched finite automaton (CEFA, see Section~\ref{sec:automaton} for the definition) $\aut_1$,  as illustrated in Figure~\ref{fig:overview}(a). 
%Then we obtain a cost-enriched finite automaton (CEFA) $\aut_1$ that is illustrated in Figure~\ref{fig:overview}(a), where the self-loops around $p_0, p_1, p_2$ represent the first, second, and third counting operator respectively, in which $r_1, r_2, r_3$ are incremented. 
%Moreover, $r_0$ is incremented in all the three self-loops. 
Furthermore, the counting bounds are specified by the accepting condition $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$. 
A string $w$ is accepted by $\aut_1$ if, when reading the characters in $w$, $\aut_1$ applies the transitions to update the state and the values of registers, and reaches a final state $q$ in the end, moreover, the resulting values of the three registers, say $v_1, v_2, v_3$, satisfy the accepting condition. 
Note that we represent the counting operators symbolically by registers, instead of unfolding them explicitly. 
Moreover, we construct another CEFA $\aut_2$ for $\Sigma^* c^+$ (see Figure~\ref{fig:overview}(b)), which is actually an NFA, since $\Sigma^* c^+$ contains no (explicit) counting operators. Finally, we construct the third CEFA $\aut_3$ (see Figure~\ref{fig:overview}(c)), where a register $r_4$ is used to denote the length of strings and the accepting condition is $\ltrue$. (See Section~\ref{subsec:regex2cefa} for more details about the construction of CEFA.)
%where the register $r_0$ is used to record the length of strings.

@@

We determinize $\cB$, apply the minimization algorithm to the resulting deterministic finite automaton (DFA), and obtain a DFA $\cC$, as illustrated in Figure~\ref{fig:overview:product:reduced}(b). 
Note that $\cC$ contains only three states $q''_0, q''_1, q''_2$ and six transitions, while $\cB$ contains four states and eight transitions. 
Although the determinization and minimization may not reduce the sizes of NFA in the worst case, the experimental results in Section~\ref{sec:implementation} show us that they do reduce the sizes of NFA constructed from the regular expressions in practice. 
%thus containing much less number of transitions than $\aut_1 \times \aut_2$\footnote{This is the case even if we consider subsets of characters, e.g. $\{a\}, \{b\}, \{c\}, \Sigma \setminus \{a, b, c\}$, in the transitions of $\aut_1 \times \aut_2$.}.
%Then we reinterpret $\cC$ as a CEFA with the acceptance condition $\alpha(r_0, r_1, r_2, r_3)$. 
%Finally, we compute an LIA formula $\psi'(i_1, \cdots, i_6)$ from $\cC$ and reduce the satisfiability of the string constraint in~(\ref{eqn-running}) to the satisfiability of the following existential LIA formula
%\begin{equation}\label{eqn-LIA-reduced}
%\psi'(i_1, \cdots, i_6) \wedge \bigwedge \limits_{0 \le j \le 3} r_j = \sum \limits_{k \in [6]}  i_k v'_{k, j+1} \wedge \alpha(r_0, r_1, r_2, r_3),
%\end{equation}
%where $\myvec{v'_{k}}=(v'_{k,1}, v'_{k,2}, v'_{k,3}, v'_{k,4})$ is the vector of updates of registers of the transition $\tau'_k$ for $k \in [6]$.

@@

At last, for the string constraints that are satisfied by a collection of strings whose lengths are small, computing LIA formulas from NFA and solving them by SMT solvers are still expensive.  
Therefore, we introduce a heuristic of searching for solutions based on \emph{under approximations}, before computing LIA formulas and solving them by SMT solvers. More precisely, we enforce a constant upper bound $B$ on the lengths of strings and search for a $B$-bounded satisfiable assignment, that is, a satisfiable assignment where the lengths of strings are bound by $B$. Initially, we set $\ell$, an integer variable to denote the lengths of strings, to $1$. If there does not exist an $\ell$-bounded satisfiable assignment, then we let $\ell:=\ell+1$ and continue the search, until finding a satisfiable assignment or reaching the bound $B$. For instance, if we remove $|x| > 120$ from the string constraint~(\ref{eqn-running}), then the resulting constraint is satisfiable. Setting the bound $B$ to $3$ is sufficient to find a solution, which is witnessed by the following accepting run of $\cC$: $q^{\prime\prime}_0 \xrightarrow{(1,0,0,1)} q^{\prime\prime}_0 \xrightarrow{(0,1,0,1)} q^{\prime\prime}_1 \xrightarrow{(0,0,0,1)} q^{\prime\prime}_2$. Note that the accepting run produces $(1,1,0,3)$, a vector of register values,  which satisfies $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$. (See Section~\ref{subsec:cefadec} for more details of the heuristic.)

>>>>

We first construct an automaton for the regular expression $(\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}}$. We introduce three registers, say $r_1, r_2, r_3$, to represent the three counting operators.  
%Moreover, we introduce another register, say $r_0$, to denote the length of strings. 
Then we construct a nondeterministic finite automaton (NFA) for $(\Sigma \setminus a)^* (\Sigma \setminus b)^* (\Sigma \setminus c)^*$, and add the updates of registers (e.g. $r_1++$ for incrementing $r_1$) to transitions, resulting in a so-called cost-enriched finite automaton (CEFA, see Section~\ref{sec:automaton} for the definition) $\aut_1$,  as illustrated in Figure~\ref{fig:overview}(a). 
%Then we obtain a cost-enriched finite automaton (CEFA) $\aut_1$ that is illustrated in Figure~\ref{fig:overview}(a), where the self-loops around $p_0, p_1, p_2$ represent the first, second, and third counting operator respectively, in which $r_1, r_2, r_3$ are incremented. 
%Moreover, $r_0$ is incremented in all the three self-loops. 
Furthermore, the counting bounds are specified by the accepting condition $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$. 
A string $w$ is accepted by $\aut_1$ if, when reading the characters in $w$, $\aut_1$ applies the transitions to update the state and the values of registers, and reaches a final state $q$ in the end, moreover, the resulting values of the three registers, say $v_1, v_2, v_3$, satisfy the accepting condition. 
Note that we represent the counting operators symbolically by registers, instead of unfolding them explicitly. 
Moreover, we construct another CEFA $\aut_2$ for $\Sigma^* c^+$ (see Figure~\ref{fig:overview}(b)), which is actually an NFA, since $\Sigma^* c^+$ contains no counting operators (Kleene plus is not seen as a counting operator here). Finally, we construct the third CEFA $\aut_3$ (see Figure~\ref{fig:overview}(c)), where a register $r_4$ is used to denote the length of strings and the accepting condition is $\ltrue$. (See Section~\ref{subsec:regex2cefa} for more details about the construction of CEFA.)
%where the register $r_0$ is used to record the length of strings.

@@

We determinize $\cB$, apply the minimization algorithm to the resulting deterministic finite automaton (DFA), and obtain a DFA $\cC$, as illustrated in Figure~\ref{fig:overview:product:reduced}(b). 
Note that $\cC$ contains only three states $q''_0, q''_1, q''_2$ and six transitions, while $\cB$ contains four states and eight transitions. 
Although the determinization and minimization may not reduce the sizes of NFA in the worst case, the experiment results show us that they do reduce the sizes of NFA constructed from the regular expressions in practice. (See Table~\ref{tab:results_heuristics}, where adding the size-reduction techniques reduces the running time for 46.4\%.) 
%thus containing much less number of transitions than $\aut_1 \times \aut_2$\footnote{This is the case even if we consider subsets of characters, e.g. $\{a\}, \{b\}, \{c\}, \Sigma \setminus \{a, b, c\}$, in the transitions of $\aut_1 \times \aut_2$.}.
%Then we reinterpret $\cC$ as a CEFA with the acceptance condition $\alpha(r_0, r_1, r_2, r_3)$. 
%Finally, we compute an LIA formula $\psi'(i_1, \cdots, i_6)$ from $\cC$ and reduce the satisfiability of the string constraint in~(\ref{eqn-running}) to the satisfiability of the following existential LIA formula
%\begin{equation}\label{eqn-LIA-reduced}
%\psi'(i_1, \cdots, i_6) \wedge \bigwedge \limits_{0 \le j \le 3} r_j = \sum \limits_{k \in [6]}  i_k v'_{k, j+1} \wedge \alpha(r_0, r_1, r_2, r_3),
%\end{equation}
%where $\myvec{v'_{k}}=(v'_{k,1}, v'_{k,2}, v'_{k,3}, v'_{k,4})$ is the vector of updates of registers of the transition $\tau'_k$ for $k \in [6]$.

@@

At last, for the string constraints that are satisfied by a collection of strings whose lengths are small, computing LIA formulas from NFA and solving them by SMT solvers are still expensive.  
Therefore, we introduce a heuristic of searching for solutions based on \emph{under approximations}, before computing LIA formulas and solving them by SMT solvers. More precisely, we enforce a constant upper bound $B$ on the lengths of strings and search for a $B$-bounded satisfiable assignment, that is, a satisfiable assignment where the lengths of strings are bound by $B$. Initially, we set $\ell$, an integer variable to denote the lengths of strings, to $1$. If there does not exist an $\ell$-bounded satisfiable assignment, then we let $\ell:=\ell+1$ and continue the search, until finding a satisfiable assignment or reaching the bound $B$. For instance, if we remove $|x| > 120$ from the string constraint~(\ref{eqn-running}), then the resulting constraint is satisfiable. Setting the bound $B$ to $3$ is sufficient to find a solution for the string constraint in~(\ref{eqn-running}), which is witnessed by the following accepting run of $\cC$: $q^{\prime\prime}_0 \xrightarrow{(1,0,0,1)} q^{\prime\prime}_0 \xrightarrow{(0,1,0,1)} q^{\prime\prime}_1 \xrightarrow{(0,0,0,1)} q^{\prime\prime}_2$. Note that the accepting run produces $(1,1,0,3)$, a vector of register values,  which satisfies $1 \le r_1 \le 60 \wedge 1 \le r_2 \le 60 \wedge 0 \le r_3 \le 60$. (See Section~\ref{subsec:cefadec} for more details.)

We implement all the aforementioned techniques on top of OSTRICH, resulting into a solver $\ostrichrecl$. It turns out that $\ostrichrecl$ is able to solve the string constraint in~(\ref{eqn-running}) within one second, while the state-of-the-art strings solvers cannot solve it within 300 seconds.
